---
title: "Explore_and_Predict"
author: "Sonia Tripolitano"
date: "16 October 2017"
output: html_document
---

## Step 1 - Cross Validation; Segmenting into Training, Testing & Validation

Given that there are many observations I decide to segment the original training
data using the benchmark 60% training/ 20% testing / 20% validation method

```{r}
        setwd("C:/Users/Stephane/Documents/Sonia/Learning/coursera/3_Pract_Machine_Learning/_Assignment")
        
        origtrain<-read.csv("pml-training.csv", 
                            na.strings=c("NA","#DIV/0!"," ",""),
                            stringsAsFactors = FALSE
                                    ) 
        
        set.seed(2601)
        library(caret)
        inTrain<-createDataPartition(origtrain$classe,p=0.8,list=FALSE)
        trainingA<-origtrain[inTrain,]
        validate<-origtrain[-inTrain,]  # 20% of the original training set is validation
        
        inTrainB<-createDataPartition(trainingA$classe,p=0.75,list=FALSE)
        training<-trainingA[inTrainB,]  # 60% of the original training set is training
        testing<-trainingA[-inTrainB,]  # 20% of the original training set is testing
        
```

## Step 2 - Explore the training data

I spent A LOT of time on this. Summarising variables, plotting, getting a feel 
for the dataset. Some of the key observations were:

- The first eight columns represent id, person, time stamps and window captured
        which are unique to a particular measurement row
- Columns 9 to 159 are the metrics of interest
- Column 160 includes the predictor variable classe which has 5 levels

- Many columns had a high number of NAs
- The NAs do not appear to be missing at random because many columns have the 
        exact same number of missing values.
- The NAs appear related to the "new_window" variable being yes 

code used (against a subset of columns) is shown below

```{r}
        str(training[,1:10])
        lapply(training[,10:20],summary)
        
```


- Most of the variables are skewed in some way, some are multi-modal

```{r pressure, echo=FALSE}
        
        histogram(training$magnet_dumbbell_x,main="Histogram of magnet_dumbbell_x")
```


